Quantile Regressions via Method of Moments with multiple fixed effects \tnoteref{t1}

### 2023-11-13 article/article_old.qmd
s
s

---
title: Quantile Regressions via Method of Moments with multiple fixed effects \tnoteref{t1}
author:
  - name: Fernando Rios-Avila
    email: friosavi@levy.org
    affiliations: 
        - id: levy
          name: Levy Economics Institute of Bard College          
    attributes:
        corresponding: true
  - name: Leonardo Siles
    email: leonardosiles@hotmail.com
    affiliations:
        - name: Universidad Católica Boliviana
          id: ucb
  - name: Gustavo Canavire-Bacarreza
    email: gcanavire@worldbank.org
    affiliations:
        - id: wb
          name: World Bank                 
abstract: |
  This paper proposes a new method to estimate quantile regressions with multiple fixed effects. The method, which expands on the strategy proposed by @mss2019, allows for the inclusion of multiple fixed effects and provides various alternatives for the estimation of standard errors. We provide Monte Carlo simulations to show the finite sample properties of the proposed method in the presence of two sets of fixed effects. Finally, we apply the proposed method to estimate the determinants of the surplus of government as a share of GDP, allowing for both time and country fixed effects.

keywords: 
  - Fixed effects
  - Linear heteroskedasticity
  - Location-scale model
  - Quantile regression

@@

<!--
How to Render this?
Lik this:
quarto render article.qmd --to elsevier-pdf

-->

# Introduction

@@

A relatively recent development in the literature has focused on extending quantile regressions analysis to include individual fixed effects in the framework of panel data. However, as described in @neymanscott1948, and @lancaster2000,  when individual fixed effects are included in quantile regression analysis an incidental parameter problem is generated. While many strategies have been proposed for estimating this type of model (see @galvao2017quantile for a brief review), neither has become standard because of their restrictive assumptions in regard to the inclusion of individual and multiple fixed effects, the computational complexity, and implementation.

More recently, @mss2019 (MSS hereafter) proposed a methodology based on a conditional location-scale model, similar to the one described in @he1997 and @zhao2000, for the estimation of quantile regressions models for panel data via a method of moments. This method allows individual fixed effects to have heterogeneous effects on the entire conditional distribution of the outcome, rather than constraining their effect to be a location shift only, as in @canay2011, @koenker2004, and @lamarche2010.

In principle, under the assumption that data-generating process behind the data is based on a multiplicative heteroskedastic process that is linear in parameters [@mss2019, @he1997, @zhao2000, @cameron2005], the effect of a variable $X$ on the $q_{th}$ quantile can be derived as the combination of a location effect, and scale effect moderated by the quantile of an underlying i.i.d. error. For statistical inference, MSS derives the asymptotic distribution of the estimator, suggesting the use of bootstrap standard errors, as well.

This methodology is not meant to substitute for the use of standard quantile regression analysis. That said, given the assumptions required for the identification of the model, it provides a simple and fast alternative for the estimation of quantile regression models with individual fixed effects.

In this framework, our paper expands on @mss2019, following some of the suggestions by these authors regarding further research. First, making use of the properties of generalized method of moments (GMM) estimators, we derive various alternatives for the estimation of standard errors based on the empirical influence functions of the estimators. Second, we reconsider the application of Frisch-Waugh-Lovell (FWL) theorem [@frishwaugh1933 and @lovell1963] to extend the MSS estimator to allow for the inclusion of multiple fixed effects, for example, individual and year fixed effects.

The rest of the paper is structured as follows: section 2 presents the basic setup of the location-scale model described in @he1997 and @zhao2000, tying the relationship between the standard quantile regression model and the location-scale model. It also revisits the methodology of MSS, proposing alternative estimators for the standard errors based on the properties of GMM estimators and the empirical influence functions. It also shows that the FWL theorem can be used to control for multiple fixed effects. Section 3 presents the results of a small simulation study and section 4 illustrates the application of the proposed methods with two empirical examples. Section 5 concludes.

@@

Quantile regressions are used to identify relationships between the explanatory variables $X$ and the conditional quantiles of the dependent variable $Q_y(\tau|X)$. This relationship is commonly assumed to follow a linear functional form:

@@

This allows for nonlinearities in the effect of $X$ on $Y$ across all values of $\tau$. This formulation can also be related to a random coefficient model, in which all coefficients are assumed to be some nonlinear function of $\tau$ where $\tau$ follows a random uniform distribution.

An alternative formulation of quantile regressions is the location-scale model. This approach assumes that the conditional quantile of $Y$ given $X$ and $\tau$ can be expressed as a combination of two models: the location model, which describes the central tendency of the conditional distribution; and the scale model, which describes deviations from the central tendency:

@@

Here, the location parameters $\beta$ are typically identified using a linear regression model (as in @mss2019) or a median regression (as in @zhao2000) and the scale parameters $\gamma(\tau)$ can be estimated using standard approaches.

@@

One characteristic of this estimator is that the $\beta(\tau)$ coefficients are identified locally and thus the estimated quantile coefficients will exhibit considerable variation when analyzed across $\tau$. It is also implicit that if one requires an analysis of the entire distribution, it would be necessary to estimate the model for each quantile.[^1]

[^1]: There are other estimators that provide smoother estimates for the quantile regression coefficients using a kernel local weighted approach [@kaplan2017], as well as identifying the full set of quantile coefficients while simultaneously assuming some parametric functional forms [@frumentobotai2016].

@@

Under the assumption that $\varepsilon$ is an i.i.d. unobserved random variable that is independent of $X$, the conditional quantile of $Y$ given $X$ and $\tau$ can be written as

@@

In this setup, the traditional quantile coefficients are identified as the location model coefficients plus the scale model coefficients moderated by the $\tau_{th}$ unconditional quantile of the standardized error $\varepsilon$. For simplicity we will use $q(\tau)$ to denote $Q_\varepsilon(\tau)$ in the rest of the paper.

@@

While this specification imposes a strong assumption on the DGP, it has two advantages over the standard quantile regression model. First, because the location-scale model can be identified globally, with only a single parameter ($q(\tau)$) requiring local estimation, this estimation approach will be more efficient than the standard quantile regression model (@zhao2000). Second, under the assumption that $X\gamma$ is strictly positive, the model will produce quantile coefficients that do not cross.

Following MSS, the quantile regression model defined by @eq-eq5 can be estimated using a generalized method of moments approach. And while it is possible to identify all coefficients ($\beta,\gamma, q(\tau)$) simultaneously, we describe and use the implementation approach advocated by MSS, which identifies each set of coefficients separately.

First, the location model can be estimated using a standard linear regression model, where the dependent variable is the outcome $Y$ and the independent variables are the explanatory variables $X$ (including a constant) with an error $u$, which is by definition heteroskedastic. In this case, the location-model coefficients are identified under the following condition:

@@

Second, after the location model is estimated, the scale coefficients can be identified by modeling heteroskedasticity as a linear function of characteristics $X$. For this we use the absolute value of the errors from the location model $u$ as dependent variable, which allows us to estimate the conditional standard deviation (rather than conditional variance) of the errors. In this case, the coefficients are identified under the following condition:

@@

where one identifies the quantile of the error $\varepsilon$ using standardized errors $\frac{y_i-x_i'\beta}{x_i'\gamma}$ or by finding the values that identify the overall quantile coefficients $\beta(\tau)=\beta + \gamma q(\tau)$. Afterwards, the conditional quantile coefficients are simply defined as the combination of the location and scale coefficients.

@@

As discussed in the previous section, the estimation of quantile regression coefficients using the location-scale model with heteroskedastic linear errors can be estimated using a the following set of moments, which fits within the GMM framework:

@@

Under the conditions described in @newey_chapter_1994 (see section 7), @cameron2005 (see chapter 6.3.9), or as shown in @mss2019, the location, scale, and residual quantile coefficients are asymptotically normal.[^3]

[^3]: @zhao2000 also shows that the quantile coefficients for the location-scale model follow a normal distribution, but uses the assumption that the location model is derived using a least absolute deviation approach (median regression).

Call $\theta=[ \beta' \ \ \gamma' \ \ q(\tau) ]'$ the set of coefficients that are identified by the moment conditions in @eq-eq10, a just identified model, and the function $h_i$ is a vector function that stacks all the moments described in @eq-eq10 at the individual level. Then $\hat\theta$ follows a normal distribution with mean $\theta$ and variance-covariance matrix $V(\theta)$ that is estimated as

@@

which is equivalent to the Eicker-White heteroskedasticity-consistent estimator for least-squares estimators.

Here, the inner product is the moment covariance matrix and $\bar{G}(\theta)$ is the Jacobian matrix of the moment equations evaluated at $\hat\theta$.

@@

In this framework, the quantile regression coefficients, a combination of the location-scale-quantile estimates, will follow a normal distribution with mean $\beta(\tau) = \beta+\gamma q(\varepsilon | \tau)$ and variance-covariance matrix equal to

@@

where $\Xi$ is a $k \times (2k+1)$ matrix defined as

@@

While it is possible to estimate the variance-covariance matrix using simultaneous model estimation for a just identified model, it is more efficient to estimate each set of coefficients separately. Afterward, the variance-covariance matrix can be estimated using the empirical influence functions of the estimators (see @jann_2020 for an overview of the application and @hampel2005 for an in-depth review).

Specifically, given an arbitrary vector of empirical influence functions $\lambda_i(\theta)$, the variance-covariance matrix can be estimated as

@@

For the specific case of quantile regressions through moments, the influence functions for the location, scale, and quantile coefficients are[^4]

@@

The different types of standard errors estimation thus depend on the assumptions imposed for the estimation of $V(\theta)$.

@@

The first and most natural standard error estimator is given by equation @eq-eqvcv. This is equivalent to the Eicker-White heteroskedasticity-consistent estimator for least-squares estimators. Considering the location-scale model, the variance-covariance matrix for the quantile coefficients can be estimated as

@@

This estimator of standard errors should be robust to arbitrary heteroskedasticity. However, because the location-scale specification relies on the correct specification of the model heteroskedasticity, large differences in standard errors compared to GLS-standard errors may be an indication of misspecification of the model.

@@

Because one of the typical applications of quantile regressions is the analysis of panel data, allowing for clustered standard errors at the individual level is important. If the unobserved error $\varepsilon$ is correlated within clusters, generalized least squares (GLS) standard errors could be severely biased. The standard recommendation has been to report block-bootstrap standard errors, clustered at the individual level.

Because we have access to the influence functions, it is straightforward to estimate one-way clustered standard errors.

Let $N_G$ be the total number of clusters $g$, where $g=1\dots N_G$. The clustered variance covariance matrix is given by[^5]

@@

where $S\lambda_i(\theta)$ is the sum of the influence functions over all observations within a given cluster $g$.

@@

The standard errors proposed by MSS can be understood as an application of GLS, which will be valid as long as the model for heteroskedasticity is correctly specified.[^6] To estimate the GLS standard errors, we make use of the following property:

[^6]: As discussed in most econometric textbooks, for example @cameron2005, one approach to correct for heteroskedasticity, when the heteroskedasticity functional form is known or can be estimated, is to use weighted least squares. While feasible this approach would defeat the purpose of identifying quantile effects exploiting the heteroskedasticity of the model.

@@

Under the assumption that the model for heteroskedasticity is correctly specified, we can apply the law of iterated expectations and rewrite the variance-covariance matrix as

@@

This standard error estimator is an application of GLS that accounts for the heteroskedasticity the model uses to identify the quantile coefficients. We can apply the same principle to find the GLS standard errors for the system of location-scale and quantile coefficients. To do this, we define the following modified influence functions:

$$\begin{aligned}
\tilde \lambda_{1,i}&=\tilde \lambda_{2,i}=N (X'X)^{-1}  x_i ( x_i'\gamma) \\  
\tilde \lambda_{3,i}&=x_i' \gamma  \\
 \psi_{i,1} &= \varepsilon_i \\
 \psi_{i,2} &= \tilde \varepsilon_i -1 \\
 \psi_{i,3} &= \frac{1}{ x_i' \gamma }\frac{\tau-\mathbb{1}\big( q(\tau)  \geq \varepsilon_i  \big) }{ f_{\varepsilon}(q(\tau))} - \frac{\varepsilon_i }{\bar x_i'\gamma} 
-  q(\tau) \frac{(\tilde \varepsilon_i -1\big)}{\bar x_i'\gamma}
\end{aligned}
$$

Then, the GLS standard errors for the location-scale and quantile coefficients can be estimated as

@@

$$\begin{aligned}
\hat\Omega_{jk} &= \frac{1}{N} \sum_{i=1}^N \tilde \lambda_{i,j}(\theta) \tilde \lambda_{i,k}(\theta)' \\
\hat\sigma_{jk} &= \frac{1}{N} \sum_{i=1}^N \psi_{i,j} \psi_{i,k}
\end{aligned}
$$

This estimator of stand ard errors is equivalent to the one derived by MSS using Theorem 3.

@@

Using the setup described in the previous section, @mss2019 proposes an extension to the model proposed by @he1997 that enables the estimation of quantile regression models with panel data, allowing for the inclusion of individual fixed effects. However, as the authors suggest, the methodology can be generalized to allow for the inclusion of multiple fixed effects. This type of analysis can be useful when considering data such as employer-employee linked data [@abowed2006] or teacher-student linked data [@harrissass2011]. Or, in the most common case, allowing to control for both individual and time fixed effects.

We return to the original model and now assume there are sets of unobserved heterogeneity that are assumed to be constant across observations, if they belong to common groups. In panel data, the groups would be the individual fixed effects and the time fixed effects. Without loss of generality, we can assume that the data-generating process is as follows:

@@

where we assume $x_{i}$ vary across groups $g_1$ and $g_2$ (and thus are not collinear) and that $\delta's$ and $\zeta's$ are the location and scale effects associated with groups fixed effects.[^7]

[^7]: We could just as well consider multiple sets of fixed effects.

If the dimension of groups $g_k$ is low, this model could be estimated using a dummy inclusion approach following @sec-betas, and the standard errors obtained as discussed in @sec-se. However, if the dimensionality of $g_k$ is high, the dummy inclusion approach may not be feasible. A more feasible approach is to apply the FWL theorem and partial out the impact of the group fixed effects on the control variables $x_{i}$, and the outcome of interest $y_{i}$, using a similar approach for the identification of $\sigma(x)$. In the case of unbalanced setups with multiple groups, the estimation involves iterative processes for which various approaches have been suggested and implemented (see for example, @correia_feasible_nodate, @gaure2013, @rios2015, among others).

When applying the partialing-out approach, some modifications to the approach described in @sec-betas are needed.

First, for all dependent and independent variables in the model ($w=y,x$), we partial out the group fixed effects and obtain the centered-residualized variables:

@@

Afterward, we estimate the location model using the centered-residualized variables:[^8]

[^8]: Using centered-residualized variables allows us to include a constant in the model specification, which simplifies the derivation of the influence functions. However, as with other fixed effects models, the constant is not identified and thus should not be interpreted.

@@

Because $|\hat \nu_i|$ is the dependent variable for the scale model, we apply the partialing out and recentering to this expression ($|\hat \nu_i|^{rc}$), and use that to estimate the following model:

@@

Finally, the standardized residuals $\varepsilon_i$ can be obtained as follows:

@@

The $\tau_{th}$ quantile of the error $\varepsilon$ can be estimated as usual and the variance-covariance matrices obtained in the same way as before (see @sec-se) by using $x_{i}^{rc}$ instead of $x_{i}$ when estimating the influence functions for all estimated coefficients.

@@

To show the performance of the extended strategy, we implement a small simulation study. We consider a simple model with a single explanatory variable $x$. In contrast with MSS, we consider a two-way fixed effect structure that is more general than the panel structure. For this exercise, we consider the following data-generating process:

@@

where $\alpha_{1i}\sim \chi^2(1)$, $\alpha_{2i} \sim \chi^2(1)$, and $x_i = 0.5 * (\chi_i +0.5(\alpha_{1i}+\alpha_{2i}))$, with $\chi_i \sim \chi^2(1)$. We only consider the case when the error term $\varepsilon_i$ is assumed to be normally distributed with mean 0 and standard deviation of 0.5. We assume that there are 50 mutually exclusive fixed effect groups for each set of fixed effects. Observations are assigned to each subgroup randomly using a uniform distribution between 1 and 50. We consider sample sizes of 500, 1000, 2000, and 4000 observations, which implies an average of 10, 20, 40, and 80 observations per group. The model is estimated using the location-scale model with heteroskedastic linear errors and we report the coefficients for the 25th and 75th quantiles. We run this excercise 5000 times. @tbl-sim1 reports the bias, simulated standard error, and mean squared error. We also report the results obtained using an adaptation bias-corrected estimator based on the split-panel jackknife estimator proposed by @dhaene2015.[^9] These results are labeled JKC.

[^9]: For the implementation, we first estimate the model using the full sample, then randomly assign each observation into one of two groups, and finally reestimate the quantile coefficients for each group. The bias-corrected estimator is then obtained as $\hat\beta(\tau)_{jkc}=2*\hat\beta(\tau)_{full}-0.5*\big(\hat\beta(\tau)_{s1}+\hat\beta(\tau)_{s1}\big)$.

Similar to the findings in MSS, we find that while the estimator presents a bias when the sample is small ($N=500$), this bias shrinks as the sample size increases. As MSS describes, the bias seems to be proportional to the sample size, or more precisely to $N/N_{g1}$ and $N/N_{g2}$, where $N_{g1}$ and $N_{g2}$ are the average number of observations per group. Interestingly, the bias-corrected estimator presents an almost 0 bias, even when the samples are small (about 10 observations per subgroup). Moreover, we find that despite the bias reduction obtained using the JKC estimator, the standard errors and mean squared error (MSE) are larger than without correction, which also replicates the results of MSS,

@@

As described in section @sec-se, GLS standard errors may not be appropriate if heteroskedasticity in the model is misspecified. This could happen if the heteroskedasticity is not linear in parameters or if there are correlations within groups. To show the performance of the GLS standard errors, we consider two additional simulations. To see the impact of heteroskedasticity misspecification, we assume the DGP is given by

@@

For the simulation, we run the excercise 5000 times, using a sample of size $N=2000$. The results from both simulations are reported in the @tbl-sim2. We report the average coefficient, simulation-based standard errors, and the standard errors using GLS, robust and clustered estimations.

@@

: Determinants of Government Surpluses: Individual Fixed Effects {#tbl-tb1}

>>>>

---
title: Quantile Regressions via Method of Moments with multiple fixed effects \tnoteref{t1}
author:
  - name: Fernando Rios-Avila
    email: friosavi@levy.org
    affiliations: 
        - id: levy
          name: Levy Economics Institute of Bard College          
    attributes:
        corresponding: true
  - name: Leonardo Siles
    email: leonardosiles@hotmail.com
    affiliations:
        - name: Universidad Católica Boliviana
          id: ucb
  - name: Gustavo Canavire-Bacarreza
    email: gcanavire@worldbank.org
    affiliations:
        - id: wb
          name: World Bank                 
abstract: |
  This paper proposes a new method to estimate quantile regressions with multiple fixed effects. The method expands on the strategy proposed by @mss2019, allowing for multiple fixed effects, and providing various alternatives for the estimation of Standard errors. We provide Monte Carlo simulations to show the finite sample properties of the proposed method in the presence of two sets of fixed effects. Finally, we apply the proposed method to estimate the determinants of the surplus of government as a share of GDP allowing for both time and country fixed effects.
keywords: 
  - Fixed effects
  - Linear heteroskedasticity
  - Location-scale model
  - Quantile regression

@@

```{=html}
<!--
How to Render this?
Lik this:
quarto render article.qmd --to elsevier-pdf

-->
```
# Introduction

@@

A relatively recent development in the literature has focused on extending quantile regressions analysis to include individual fixed effects in the framework of panel data. However, as described in @neymanscott1948, and @lancaster2000, when individual fixed effects are included in quantile regression analysis it generates an incidental parameter problem. While many strategies have been proposed for estimating this type of model (see @galvao2017quantile for a brief review), neither has become standard because of their restrictive assumptions in regards to the individual effects, the computational complexity, and implementation.

More recently, @mss2019 (MSS hereafter) proposed a methodology based on a conditional location-scale model similar to the one described in @he1997 and @zhao2000, for the estimation of quantile regressions models for panel data via a method of moments. This method allows individual fixed effects allowing to have heterogeneous effects on the entire conditional distribution of the outcome, rather constraining their effect to be a location shift only as in @canay2011, @koenker2004, and @lamarche2010.

In principle, under the assumption that data generating process behind the data is based on a multiplicative heteroskedastic process that is linear in parameters [@cameron2005, @mss2019, @he1997, @zhao2000], the effect of a variable $X$ on the $q_{th}$ quantile can be derived as the combination of a location effect, and scale effect moderated by the quantile of an underlying i.i.d. error. For statistical inference, MSS derives the asymptotic distribution of the estimator, suggesting the use of bootstrap standard errors, as well.

While this methodology is not meant to substitute the use of standard quantile regression analysis, given the assumptions required for the identification of the model, it provides a simple and fast alternative for the estimation of quantile regression models with individual fixed effects.

In this framework, our paper expands on @mss2019, following some of the suggestions by the authors regarding further research. First, making use of the properties of GMM estimators, we derive various alternatives for the estimation of standard errors based on the empirical Influence functions of the estimators. Second, we reconsider the application of Frisch--Waugh--Lovell (FWL) theorem (@frishwaugh1933 and @lovell1963) to extend the MSS estimator to allow for the inclusion of multiple fixed effects, for example, individual and year fixed effects.

The rest of the paper is restructured as follows. Section 2 presents the basic setup of the location-Scale model described in @he1997 and @zhao2000, tying the relationship between the standard quantile regression model, and the location and scale model. It also revisits MSS methodology, proposing alternative estimators for the standard errors based on the properties of GMM estimators and the empirical influence functions. It also shows that FWL theorem can be used to control for multiple fixed effects. Section 3 presents the results of a small simulation study and Section 4 illustrates the application of the proposed methods with one empirical example. Seccion 5 concludes.

@@

Quantile regressions are used to identify relationships between the explanatory variables $x$ and the conditional quantiles of the dependent variable $Q_y(\tau|X)$. This relationship is commonly assumed to follow a linear functional form:

@@

This allows for the effect of $X$ on $Y$ to vary across all $\tau$. This formulation can also be related to a random coefficient model, where all coefficients are assumed to be some nonlinear function of $\tau$, where $\tau$ follows a random uniform distribution.

An alternative formulation of quantile regressions is the location-scale model. This approach assumes that the conditional quantile of $Y$ given $X$ and $\tau$ can be expressed as a combination of two models: the location model, which describes the central tendency of the conditional distribution, and the scale model, which describes deviations from the central tendency:

@@

Here, the location parameters $\beta$ are typically identified using a linear regression model (as in @mss2019), or a median regression (as in @zhao2000), and the scale parameters $\gamma(\tau)$ can be estimated using standard approaches.

@@

One characteristic of this estimator is that the $\beta(\tau)$ coefficients are identified locally, and thus the estimated quantile coefficients will exhibit considerable variation when analyzed across $\tau$. It is also implicit that if one requires an analysis of the entire distribution, it would be necessary to estimate the model for each quantile.[^1]

[^1]: There are other estimators that provide smoother estimates for the quantile regression coefficients using a kernel local weighted approach [@kaplan2017], as well as identifying the full set of quantile coefficients simultaneously assuming some parametric functional forms [@frumentobotai2016].

@@

Under the assumption that $\varepsilon$ is an independent and identically distributed (iid) unobserved random variable that is independent of $X$, the conditional quantile of $Y$ given $X$ and $\tau$ can be written as:

@@

In this setup, the traditional quantile coefficients are identified as the location model coefficients, plus the scale model coefficients moderated by the $\tau_{th}$ unconditional quantile of the standardized error $\varepsilon$. For simplicity we will use $q(\tau)$ to denote $Q_\varepsilon(\tau)$ in the rest of the paper.

@@

While this specification imposes a strong assumption on the DGP, it has two advantages over the standard quantile regression model. First, because the location and scale model can be identified globally, with only a single paramater ($q(\tau)$) requiring local estimation, this estimation approach would be more efficient than the standard quantile regression model [@zhao2000]. Second, under the assumption that $X\gamma$ is strictly possitive, the model would produce quantile coefficients that do not cross.

Following MSS, the quantile regression model defined by @eq-eq5 can be estimated using a method of moments approach. And while its possible to identify all coefficients ($\beta,\gamma, q(\tau)$) simultaneously, we describe and use the implementation approach advocated by MSS which identifies each set of coefficients separately.

First, the location model can be estimated using a standard linear regression model, where the dependent variable is the outcome $y$, and the independent variables are the explanatory variables $X$ (including a constant) with an error $u$, which is by definition heteroskedastic. In this case, the location model coefficients are identified under the following condition:

@@

Second, after the location model is estimated, the scale coefficients can be identified by modeling heteroskedasticity as a linear function of characteristics $X$. For this we use the absolute value of the errors from the location model $u$ as dependent variable, which would allow us to estimate the conditional standard deviation (rather than conditional variance) of the errors. In this case, the coefficients are identified under the following condition:

@@

Where one identifies the quantile of the error $\varepsilon$ using standardized errors $\frac{y_i-x_i'\beta}{x_i'\gamma}$, or by finding the values that identify the overall quantile coefficients $\beta(\tau)=\beta + \gamma q(\tau)$. Afterwords, the conditional quantile coefficients is simply defined as the combination of the location and scale coefficients.

@@

As discussed in the previous section, the estimation of quantile regression coefficients using the location-scale model with heteroskedstic linear errors can be estimated using a the following set of moments, which fits in the Generalized Method of Moments framework:

@@

Under the conditions described in @newey_chapter_1994 (see section 7), @cameron2005 (see chapter 6.3.9) or as shown in @mss2019, the location, scale and residual quantile coefficients are asymptotically normal.[^3]

[^3]: @zhao2000 also shows that the quantile coefficients for the location-scale model also follows a normal distribution, but uses the assumption that the location model is derived using a least absolute deviation approach (median regression).

Call $\theta=[ \beta' \ \ \gamma' \ \ q(\tau) ]'$ the set of coefficients that are identified by the modement conditions in @eq-eq10, a just identified model. And the function $h_i$ is a vector function that stacks all the moments at the individual level described in @eq-eq10. Then $\hat\theta$ follows a normal distribution with mean $\theta$ and variance-covariance matrix $V(\theta)$ that is estimated as:

@@

Which is equivalent to the Eicker-White Heteroskedastic-Consistent estimator for least-squares estimators.

Here, the inner product is the moment covariance matrix, and $\bar{G}(\theta)$ is the Jacobian matrix of the moment equations evaluated at $\hat\theta$.

@@

In this framework, the quantile regression coefficients, a combination of the location-scale-quantile estimates, will follow a normal distribution with mean $\beta(\tau) = \beta+\gamma q(\tau)$ and variance-covariance matrix equal to:

@@

where $\Xi$ is a $k \times (2k+1)$ matrix defined as:

@@

While it is possible to estimate the variance-covariance matrix using simultaneous model estimation, for a just identified model, it is more efficient to estimate each set of coefficients separately. Afterwards, the variance-covariance matrix can be estimated using the empirical influence functions of the estimators (see @jann_2020 for an overview of the application, and @hampel2005 for an in-depth review).

Specifically, given an arbitrary vector of empirical influence functions $\lambda_i(\theta)$, the variance-covariance matrix can be estimated as:

@@

For the specific case of quantile regressions via momoments, the influence functions for the location, scale and quantile coefficients are:[^4]

@@

The different types of Standard errors estimation, thus, depend on the assumptions imposed for the estimation of $V(\theta)$.

@@

The first, and most natural standard error estimator is given by equation @eq-eqvcv. This is equivalent to the Eicker-White Heteroskedastic-Consistent estimator for least-squares estimators. Considering the location-scale model, the variance-covariance matrix for the quantile coefficients can be estimated as:

@@

This estimator of Standard errors should be robust to arbitrary heteroskedasticity. However, because the location-scale specification relies on the correct specification of the model heteroskedasticity, large differences in Standard errors compared to GLS-standard errors may be an indication of misspecification of the model.

@@

Because one of the typical applications of quantile regressions is the analysis of panel data, allowing for clustered standard errors at the individual level is important. If the unobserved error $\varepsilon$ is correlated within clusters, GLS-Standard errors could be severily biased. The standard recommendation has been to report block-bootstrap standard errors, clustering at the individual level.

Since we have access to the influence functions, it is straight forward to estimate one-way clustered standard errors.

Call $N_G$ to be the total number of clusteres g, where $g=1\dots N_G$. The clustered variance covariance matrix is given by:[^5]

@@

Where $S\lambda_i(\theta)$ is the sum of the influence functions over all observations within a given cluster $g$.

@@

The standard errors proposed by MSS can be understood as an application of generalized least squares (GLS), which will be valid as long as the model for heteroskedasticity is correctly specified.[^6] To estimate the GLS-Standard errors, we make use of the following property:

[^6]: As discussed in most econometric textbooks, like @cameron2005, one approach to correct for heteroskedasticity, when the heteroskedasticity functional form is known, or can be estimated, is to use weighted least squares. While feasible, however, this approach would defeat the purpose of identifying quantile effects exploiting the heteroskedasticity of the model.

@@

Under the assumption that the model for heteroskedasticity is correctly specified, we can apply the law of iterated expectations and rewrite the variance-covariance matrix as:

@@

This standard error estimator is an application of GLS that accounts for the heteroskedasticity the model uses to identify the quantile coefficients. We can apply the same principle to find the GLS-Standard errors for the system of location-scale and quantile coefficients. To do this, define the following modified influence functions:

$$\begin{aligned}
\tilde \lambda_{1,i}&=\tilde \lambda_{2,i}=N (X'X)^{-1}  x_i ( x_i'\gamma) \\  
\tilde \lambda_{3,i}&=x_i' \gamma  \\
 \psi_1 &= \varepsilon_i \\
 \psi_2 &= \tilde \varepsilon_i -1 \\
 \psi_3 &= \frac{1}{ x_i' \gamma }\frac{\tau-\mathbb{1}\big( q(\tau)  \geq \varepsilon_i  \big) }{ f_{\varepsilon}(q(\tau))} - \frac{\varepsilon_i }{\bar x_i'\gamma} 
-  q(\tau) \frac{(\tilde \varepsilon_i -1\big)}{\bar x_i'\gamma}
\end{aligned}
$$

Then, the GLS-Standard errors for the location-scale and quantile coefficients can be estimated as:

@@

$$\begin{aligned}
\hat\Omega_{ij} &= \frac{1}{N} \sum_i^N \tilde \lambda_{i}(\theta) \tilde \lambda_{j}(\theta)' \\
\hat\sigma_{ij} &= \frac{1}{N} \sum_i^N \psi_{i} \psi_{j}
\end{aligned}
$$

This estimator of Standard errors is equivalent to the one derived by MSS by Theorem 3.

@@

Using the setup described in the previous section, @mss2019 proposes an extension to the model proposed by @he1997 that would allow for the estimation of quantile regression models with panel data, allowing for the inclusion of individual fixed effects. However, as the authors suggest, the methodology can be generalized to allow for the inclusion of multiple fixed effects. This type of analysis may be useful when considering data such as employer-employee linked data [@abowed2006], or teacher-student linked data [@harrissass2011]. Or, in the most common case, allowing to control for both individual and time fixed effects.

Reconsider the original model, and assume there are sets of unobserved heterogeneity that are assumed to be constant across observations, if they belong to common groups. In panel data, the groups would be the individual fixed effects and the time fixed effects. Without loss of generality, we can assume that the data generating process is as follows:

@@

where we assume $x_{i}$ vary across groups $g_1$ and $g_2$, thus are not collinear, and that $\delta's$ and $\zeta's$ are the location and scale effects associated with groups fixed effects.[^7]

[^7]: We could just as well consider multiple sets of fixed effects

If the dimension of groups $g_k$ is low, this model could be estimated using a dummy inclussion approach following @sec-betas, and standard errors obtained as discussed in @sec-se. However, if the dimensionality of $g_k$ is high, the dummy inclusion approach may not be feasible. Instead, a more feasible approach is to apply the Frisch-Waugh-Lovell (FWL) theorem, and partial out the impact of the group fixed effects on the control variables $x_{i}$, the outcome of interest $y_{i}$, with a similar approach for the identification of $\sigma(x)$. In the case of unbalanced setups, with multiple groups, the estimation involves iterative processes for which various approaches have been suggested and implemented (see @correia_feasible_nodate, @gaure2013, @rios2015, among others).

When applying the partialing out approach, some modifications to the approach described in @sec-betas are needed.

First, for all dependent and independent variables in the model ($w=y,x$), we partial out the group fixed effects, and obtain the centered-residualized variables:

@@

Afterwards, we estimate the location model using the centered-residualized variables:[^8]

[^8]: Using centered-residualized variables allow us to include a constant in the model specification, which simplifies the derivation of the influence functions. However, as with other fixed effects models, the constant is not identified, and thus should not be interpreted.

@@

Since $|\hat \nu_i|$ is the dependent variable for the scale model, we apply the partialling out and recentering to this expression ($|\hat \nu_i|^{rc}$), and use that to estimate the model:

@@

Finally the standardized residuals $\varepsilon_i$ can be obtained as follows

@@

The $\tau_{th}$ quantile of the error $\varepsilon$ can be estimated as usual, and the variance-covariance matrices obtained in the same way as before (see @sec-se), but using $x_{i}^{rc}$ instead of $x_{i}$ when estimating the influence functions for all estimated coefficients.

@@

To show the performance of the extended strategy, we implement a small simulation study. We consider a simple model with a single explanatory variable $x$. In contrast with MSS, we consider a two-way fixed effect structure that is more general than the panel structure. For this excercise, we consider the following data generating process:

@@

where $\alpha_{1i}\sim \chi^2(1)$, $\alpha_{2i} \sim \chi^2(1)$ and $x_i = 0.5 * (\chi_i +0.5(\alpha_{1i}+\alpha_{2i}))$ with $\chi_i \sim \chi^2(1)$. We only consider the case when the error term $\varepsilon_i$ is assumed to be normally distributed with mean zero and standard deviation of 0.5. We assume that there are 50 mutually exclusive fixed effect groups for each set of fixed effects. Observations are assigned to each subgroup randomly, using a uniform distribution between 1 and 50. We consider a sample sizes of 500, 1000, 2000 and 4000 observations, which implies an average of 10, 20, 40 and 80 observations per group. The model is estimated using the location-scale model with heteroskedastic linear errors, and we report the coefficients for the 25th and 75th quantiles. We run this excercise 5000 times. @tbl-sim1 reports the bias, simulated standard error and mean squared error. We also report the results obtained using an adaptation bias-corrected estimator based on the split-panel jackknife estimator proposed by @dhaene2015.[^9] These results are labeled JKC.

[^9]: For the implementation, we estimate the model using the full sample, then, randomly assign each observation into one of two groups, and re-estimate the quantile coefficients for each group. The bias-corrected estimator is then obtained as $\hat\beta(\tau)_{jkc}=2*\hat\beta(\tau)_{full}-0.5*\big(\hat\beta(\tau)_{s1}+\hat\beta(\tau)_{s1}\big)$.

Similar to the findings in MSS, we find that the estimator presents a bias when the sample is small ($N=500$), but this bias shrinks as the sample sizes increases. As MSS describes, the bias seems to be proportional to the sample size, or more precisely to $N/N_{g1}$ and $N/N_{g2}$, where $N_{g1}$ and $N_{g2}$ are the average number of observations per group. Interestingly, the bias-corrected estimator presents an almost zero bias, even when the samples are small (about 10 observations per sub-group). Also replicating the results of MSS, despite bias reduction obtained using the JKC estimator, the standard errors and mean squared error (MSE) are larger than without correction.

@@

As described in section @sec-se, GLS-Standard errors may not be appropriate if heteroskedasticity in the model is misspecified. This could happen if the heteroskedasticity is not linear in parameters, or if there is correlations within groups. To show the performance of the GLS-Standard errors, we consider two additional simulations. To see the impact of heteroskedasticity misspecification, we assume the DGP is given by:

@@

For the simulation, we run the excercise 5000 times, using a sample of size $N=2000$. The results from both simulations are reported in the @tbl-sim2. We report the average coefficient, simulation based Standard errors, and the standard errors using GLS, robust and clustered estimations.

@@

: The determinants of government surpluses: Individual Fixed effects {#tbl-tb1}
